Statistical Stopping Criteria for Automated Screening in Systematic Reviews

### 2020-09-01 manuscript/main.tex
re-resubmission

			\title{Robust Statistical Stopping Criteria for Automated Screening in Systematic Reviews}

@@

			\begin{abstract} % abstract
				%\parttitle{First part title} %if any
				Active learning for systematic review screening promises to reduce the human effort required to identify relevant documents for a systematic review. 
				Machines and humans  work together, with humans providing training data, and the machine optimising the documents that the humans screen. This enables the identification of all relevant documents after viewing only a fraction of the total documents. 
				However, current approaches lack robust stopping criteria, so that reviewers do not know when they have seen all or a certain proportion of relevant documents. This means that such systems are hard to implement in live reviews. 
				This paper introduces a workflow with robust and flexible statistical stopping criteria, which offer real work reductions on the basis of a given confidence level of reaching a given recall.
				The stopping criteria are shown on test datasets to achieve a reliable level of recall, while still providing work reductions of on average 17\%. Other methods proposed previously are shown to provide inconsistent recall and work reductions across datasets.

@@

	This study proposes a system for estimating the recall based on random sampling of remaining documents. 
	We use a simple statistical method to iteratively test a null hypothesis that the recall achieved is less than a given target recall. If the hypothesis can be rejected, we conclude that the recall target has been achieved with a given confidence level and screening can be stopped.
	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``A recall of more than 95\% was achieved with a confidence of more than 95\%''.

@@

	However, their method does not quantify uncertainty, but can only claim that the method ``\textit{tends} to result in a conservative estimate of recall'' (emphasis ours). This is not guaranteed by the criterion itself but rather a finding of the simulation with example datasets. Further the authors do not give sufficient information to reproduce their results, providing neither code (they describe their own proprietary software), nor an equation for $\hat{p}$. Additionally, the criterion requires a tuning parameter $\delta$, which users may have insufficient information to set optimally.
	Lastly, because screening is a form of sampling without replacement, the negative hypergeometric distribution should be preferred to the negative binomial, even though the latter can be a good approximation for cases with large numbers of documents.

	These last examples are promising developments, but they all fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount, and the results are independent of model performance. In the following, we explain our own method, which provides clearly communicable estimates of recall, which manage uncertainty in a way robust to model performance.

@@

@@

	This equation is derived by combining Eqs.~\ref{eq:recall-def} and \ref{eq:rho_seen}. Because $k$ can only take integer values, $K_{tar}$ is the smallest integer that satisfies the inequality in Eq.~\ref{eq:null_hypothesis}.
	With $K_{tar}$, we can reformulate our null hypothesis: the true number of relevant documents in the sample is greater than our hypothetical value.

	\begin{equation}
	H_0 : K > K_{tar}
	\end{equation}

@@

	We can reject our null hypothesis and stop screening if the maximum probability of obtaining our observed results given our null hypothesis $p$ is below $1-\alpha$. This means, we can report the likelihood that we achieve a recall above our target as being more than $\alpha$ \footnote{The notebook, \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteria_theory.ipynb}, in the github repository accompanying this paper contains a step by step explanation of this method with code and examples}.

@@

	We calculate $p$ as above for subsets of the already screened documents. Concretely, we use subsets of documents $A_i$ by looking back to the last $i$ documents, $A_i = \{d_{N_{seen} - 1}, ..., d_{N_{seen} - i}\}$, where the documents $d$ are indexed in the order in which they have been screened. For a specific $i$, this corresponds to  random sampling beginning after seeing $i$ documents in the section above.
	Thus, we set 
	$N_{AL}$ to $i$, 
	$n$ to $N_{seen}-i$, 
	$\rho_{AL}$ to the number of relevant documents seen when $i$ documents had been seen,
	and $k$ to the number of relevant documents seen since $i$ documents had been seen, and calculate $p$ according to Eq.~\ref{eq:p-value}.
	We compute $p$ for all sets $A_i$ with $i \in {N_{seen}-1 \dots 1}$.
	This gives us a vector $\bm{p}$, representing the values of $p$ which would have been estimated at each point at which we could have stopped active learning and began random sampling. The lowest probability of our null hypothesis being true that we would have thereby obtained is given by $p_{min}$. With the vectorized implementation included in our accompanying code, these calculations are completed in less than the time it would take a human to code the next document.

@@

	Figure \ref{recall-wss} shows the actual recall and work savings achieved when each stopping criteria has been satisfied. 
	For comparison, we also include the results that would have been achieved with \textit{a priori} knowledge of the data, that is, the work saved when the 95\% recall target was actually reached. In a live systematic review, reviewers would never know when this had been reached, but these are the work savings most often reported in machine learning for systematic review screening studies.

@@

	Both the random sampling and the ranked sampling criteria achieve the target threshold of 95\% in more than 95\% of cases. 
	That this is greater than 95\% is accounted for by the fact that random sampling sometimes begins after the target has been achieved, in which case the null hypothesis would be \text{a priori} impossible.
	The ranked quasi-sampling criterion  outperforms the random sampling criterion with respect to both recall and work savings, saving a mean of 17\% of the work compared to 15\%, and missing the target in only 0.95\% compared to 3.29\% of cases. In theory, the ranked sampling criteria is conservative if the assumption holds that documents chosen by machine learning are not less likely to be relevant than those chosen at random. Based on our experiments, this assumption seems reasonable, and accounts for the higher recall. Because the ranked quasi-sampling criterion can flexibly choose its sample, whereas the random criterion has to wait for a random sample to be triggered, the criterion is also triggered earlier, as it can make use of more data. This accounts for the higher work savings.

	The baseline sampling criteria (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%. This is in line with our expectations that, due to random sampling error, the expected number of documents will often be over-estimated or under-estimated, resulting in zero work savings or poor recall.

@@

	Figure \ref{H0paths} shows the recall and the probability of the null hypothesis for the best performing iteration of four datasets. Although the 95\% recall target is achieved very quickly in the Radjenovic dataset, the null hypothesis cannot be excluded until much later. This is because the dataset has only 47 relevant documents out of a population of 5,999. After the 95\% recall target was achieved, 45 out of 47 relevant documents had been seen and 5,029 documents remained. The null hypothesis was therefore that 3 or more of these 5,029 documents were relevant, which requires a lot of evidence to disprove. The burden of proof was smaller in the case of the Proton Beam dataset: at the point that the 95\% recall threshold was reached, the null hypothesis to disprove was that a minimum of 13 out of 3,369 remaining documents were relevant.

@@

	Different use cases for machine learning enhanced screening may also carry different requirements for recall, or different tolerances for uncertainty. These can be flexibly accommodated within our stopping criterion. Importantly, the ability to make probabilistic statements about the chance of achieving a given recall target makes it possible to clearly communicate the implications of using machine learning enhanced screening to readers and reviewers who are not machine learning specialists. This is extremely important in live systematic reviews.

@@

	One caveat in the derivation of our criteria is that we did not address the potential problem of multiple testing formally. Such a derivation is mathematically challenging and beyond the scope of this paper. However, the performance of the criteria shows that this is of limited practical concern. Formally describing screening procedures with iterative testing should be a next step towards even more rigorous stopping criteria and should be fully worked out in future research.

@@

	This paper demonstrates the drawbacks of existing stopping criteria for machine learning approaches to document screening, particularly with regard to reliability. We propose a simple method that delivers reliable recall, independent of machine learning approach or model performance. Our robust statistical stopping criteria allow users to easily communicate the implications of their use of machine learning, making machine learning enhanced screening ready for live reviews.

>>>>

			\title{Statistical Stopping Criteria for Automated Screening in Systematic Reviews}

@@

			\begin{abstract} % abstract
				%\parttitle{First part title} %if any
				Active learning for systematic review screening promises to reduce the human effort required to identify relevant documents for a systematic review. 
				Machines and humans  work together, with humans providing training data, and the machine optimising the documents that the humans screen. This enables the identification of all relevant documents after viewing only a fraction of the total documents. 
				However, current approaches lack robust stopping criteria, so that reviewers do not know when they have seen all or a certain proportion of relevant documents. This means that such systems are hard to implement in live reviews. 
				This paper introduces a workflow with flexible statistical stopping criteria, which offer real work reductions on the basis of rejecting a hypothesis of having missed a given recall target with a given level of confidence.
				The stopping criteria are shown on test datasets to achieve a reliable level of recall, while still providing work reductions of on average 17\%. Other methods proposed previously are shown to provide inconsistent recall and work reductions across datasets.

@@

	This study proposes a system for estimating the recall based on random sampling of remaining documents. 
	We use a simple statistical method to iteratively test a null hypothesis that the recall achieved is less than a given target recall. If the hypothesis can be rejected, we conclude that the recall target has been achieved with a given confidence level and screening can be stopped.
	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``We reject the null hypothesis that we achieve a recall of less than 95\% with a significance level of 5\%''.

@@

	However, their method does not quantify uncertainty, but can only claim that the method ``\textit{tends} to result in a conservative estimate of recall'' (emphasis ours). This is not guaranteed by the criterion itself but rather a finding of the simulation with example datasets. Further, the authors do not give sufficient information to reproduce their results, providing neither code (they describe their own proprietary software), nor an equation for $\hat{p}$. Additionally, the criterion requires a tuning parameter $\delta$, which users may have insufficient information to set optimally.
	Lastly, because screening is a form of sampling without replacement, the negative hypergeometric distribution should be preferred to the negative binomial, even though the latter can be a good approximation for cases with large numbers of documents.

	These last examples are promising developments, but they all fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount, and the results are independent of model performance. In the following, we explain our own method, which provides clearly communicable estimates of recall, and which manage uncertainty in a way robust to model performance.

@@

	Accordingly, the alternative hypothesis is that recall is equal to or greater than our target:

	\begin{equation}
	H_1 : \tau \geq \tau_{tar}
	\end{equation}

@@

	This equation is derived by combining Eqs.~\ref{eq:recall-def} and \ref{eq:rho_seen}. Because $k$ can only take integer values, $K_{tar}$ is the smallest integer that satisfies the inequality in Eq.~\ref{eq:null_hypothesis}.
	With $K_{tar}$, we can reformulate our null hypothesis: the true number of relevant documents in the sample is greater than or equal to our hypothetical value.

	\begin{equation}
	H_0 : K \geq K_{tar}
	\end{equation}

@@

	We can reject our null hypothesis and stop screening if the maximum probability of obtaining our observed results given our null hypothesis $p$ is below $1-\alpha$ \footnote{The notebook, \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteria_theory.ipynb}, in the github repository accompanying this paper contains a step by step explanation of this method with code and examples}. To further investigate the accuracy of the test, we perform an experiment drawing 1 million random samples in 6 scenarios with different characteristics. We vary the value of $\rho_{AL}$ to simulate starting random sampling with different levels of recall achieved.

	\begin{figure}
		\centering
		\includegraphics[width=\linewidth]{2_figs_hyper_validation.png}
		\caption{The distribution of achieved recall values given our random sampling stopping criterion for 6 scenarios with different recall values at the start of sampling.}
		\label{fig:hyper-valid}
	\end{figure}

	Figure \ref{fig:hyper-valid} shows that in each case, as long as recall is lower than the target recall when sampling begins, the percentage of times that the criteria is triggered too early is within two tenths of a percentage point of 5\% and the the 5th percentile of achieved recall values is within two tenths of a percentage point of the target recall 95\%.

@@

	We calculate $p$ as above for subsets of the already screened documents. Concretely, we use subsets of documents $A_i$ by looking back to the last $i$ documents, $A_i = \{d_{N_{seen} - 1}, ..., d_{N_{seen} - i}\}$, where the documents $d$ are indexed in the order in which they have been screened. For a specific $i$, this corresponds to  random sampling beginning after seeing $i$ documents in the section above.
	Thus, we set 
	$N_{AL}$ to $i$, 
	$n$ to $N_{seen}-i$, 
	$\rho_{AL}$ to the number of relevant documents seen when $i$ documents had been seen,
	and $k$ to the number of relevant documents seen since $i$ documents had been seen, and calculate $p$ according to Eq.~\ref{eq:p-value}.
	We compute $p$ for all sets $A_i$ with $i \in {N_{seen}-1 \dots 1}$.
	This gives us a vector $\bm{p}$, representing the values of $p$ which would have been estimated at each point at which we could have stopped active learning and began random sampling. The point at which the p-value for our null hypothesis is lowest is given by $p_{min}$. With the vectorized implementation included in our accompanying code, these calculations are completed in less than the time it would take a human to code the next document.

@@

	Figure \ref{recall-wss} shows the actual recall and work savings achieved when each stopping criterion has been satisfied. 
	For comparison, we also include the results that would have been achieved with \textit{a priori} knowledge of the data, that is, the work saved when the 95\% recall target was actually reached. In a live systematic review, reviewers would never know when this had been reached, but these are the work savings most often reported in machine learning for systematic review screening studies.

@@

	Both the random sampling and the ranked sampling criteria achieve the target threshold of 95\% in more than 95\% of cases. 
	That this is greater than 95\% is accounted for by the fact that random sampling sometimes begins after the target recall has been achieved, in which case the null hypothesis would be \text{a priori} impossible.
	The ranked quasi-sampling criterion  outperforms the random sampling criterion with respect to both recall and work savings, saving a mean of 17\% of the work compared to 15\%, and missing the target in only 0.95\% compared to 3.29\% of cases. In theory, the ranked sampling criteria is conservative if the assumption holds that documents chosen by machine learning are not less likely to be relevant than those chosen at random. Based on our experiments, this assumption seems reasonable, and accounts for the higher recall. Because the ranked quasi-sampling criterion can flexibly choose its sample, whereas the random criterion has to wait for a random sample to be triggered, the criterion is also triggered earlier, as it can make use of more data. This accounts for the higher work savings.

	The baseline sampling criteron (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%. This is in line with our expectations that, due to random sampling error, the expected number of documents will often be over-estimated or under-estimated, resulting in zero work savings or poor recall.

@@

	Figure \ref{H0paths} shows the recall and the p-value for the null hypothesis for the the iteration where the recall target is reached first for four datasets. Although the 95\% recall target is achieved very quickly in the Radjenovic dataset, the null hypothesis cannot be excluded until much later. This is because the dataset has only 47 relevant documents out of a population of 5,999. After the 95\% recall target was achieved, 45 out of 47 relevant documents had been seen and 5,029 documents remained. The null hypothesis was therefore that 3 or more of these 5,029 documents were relevant, which requires a lot of evidence to disprove. The burden of proof was smaller in the case of the Proton Beam dataset: at the point that the 95\% recall threshold was reached, the null hypothesis to disprove was that a minimum of 13 out of 3,369 remaining documents were relevant.

@@

	Different use cases for machine learning enhanced screening may also carry different requirements for recall, or different tolerances for uncertainty. These can be flexibly accommodated within our stopping criterion. Importantly, the ability to make statements about the authors' confidence in achieving a given recall target makes it possible to clearly communicate the implications of using machine learning enhanced screening to readers and reviewers who are not machine learning specialists. This is extremely important in live systematic reviews.

@@

	One caveat in the derivation of our criteria is that we did not address the problem of multiple testing formally. Such a derivation is mathematically challenging and beyond the scope of this paper. However, the performance of the criteria shows that this is of limited practical concern. Formally describing screening procedures with iterative testing should be a next step towards even more rigorous stopping criteria and should be fully worked out in future research.

@@

	This paper demonstrates the drawbacks of existing stopping criteria for machine learning approaches to document screening, particularly with regard to reliability. We propose a simple method that delivers reliable recall, independent of machine learning approach or model performance. Our statistical stopping criteria allow users to easily communicate the implications of their use of machine learning, making machine learning enhanced screening ready for live reviews.
